---
title: "Basic sensor data wrangling"
author: "Amelia Ritger"
date: "11/9/2021"
output: html_document
---

# Load packages
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      messages = FALSE, 
                      error = FALSE,
                      warning = FALSE)

library(tidyverse)
library(janitor)
library(here)
library(lubridate)
library(scales)
library(rvest) #to scrape tide data from the internet
library(reshape) #to ultimately merge BML, LOL, and ALG
library(gt) #create nice table for presentations
library(webshot) #save gt() table
library(highcharter) #create highchart to interact with data
library(webshot) #save highcharts to .png files
library(plotly) #create plots with multiple y axes
library(filesstrings) #move plotly figure to proper folder location
```

## Read in the data
```{r}
all <- read_csv(here("data","sensor-data_all.csv")) %>% 
  clean_names() %>%
  mutate(date_time=ymd_hms(date_time), #apply lubridate to date/time column
         date=format(date_time, '%Y/%m/%d'), #create only date column
         time=format(date_time, '%H:%M:%S')) %>% #create only time column
 select(site, sensor_number, date_time, date, time, temp_c, p_h) %>%
    mutate(site=replace(site, site=="LOL", "Lompoc Landing"),
           site=replace(site, site=="ALG", "Alegria"),
           site=replace(site, site=="BML", "Bodega Bay")) #dplyr::rename locations
```

## Plot up the pH
```{r}
#set site order for plotting (legend)
all$site <- factor(all$site, levels=c("Alegria", "Lompoc Landing", "Bodega Bay"))

all_filter <- all %>%
  filter(p_h < 8.5) %>%
  mutate(removeit = ifelse(site=="Alegria" & date_time<ymd_hms("2021-06-30 23:00:00"), "remove", "keep")) %>%
  filter(removeit=="keep")

ggplot(all_filter, aes(x=date_time, y=p_h, group=site)) +
  geom_line(aes(color=site), size=0.7) +
  geom_point(aes(color=site), size=0.5) +
  scale_x_datetime(breaks = scales::date_breaks("1 week"), 
                    labels = date_format("%m/%d %H:%m")) +
  xlab("Date time") +
  ylab("pH") +
  theme_bw() +
  theme(legend.title=element_blank(),
        axis.text.x=element_text(angle=45, vjust = 1, hjust=1, size=12),
        axis.title.x=element_text(size=15),
        axis.text.y=element_text(size=12),
        axis.title.y=element_text(size=15),
        legend.text=element_text(siz=12))

#ggsave(here("figures", "june-sept.png"), height=20, width=40, units="cm")
```

### Plot temp up for good measure
```{r}
ggplot(all, aes(x=date_time, y=temp_c, group=site)) +
  geom_line(aes(color=site), size=0.7) +
  geom_point(aes(color=site), size=0.5) +
  scale_x_datetime(breaks = scales::date_breaks("1 week"), 
                    labels = date_format("%m/%d %H:%m")) +
  xlab("Date time") +
  theme_bw() +
  theme(axis.text.x=element_text(angle=90))

#ggsave(here("figures", "sensors","june-sept_temp.png"), height=20, width=40, units="cm")
```

# Plot each site, temp and pH 

#### Customize HTML

Sites: 

- Bodega (Bodega%20Harbor%20entrance%2C%20California)
- Lompoc (Point%20Arguello%2C%20California)
- Alegria (Gaviota%2C%20California)

Glen: number of days to extract

Interval:

 - 10 minutes (00%3A10)
 - 15 minutes (00%3A15)
 - 1 hour (01%3A00)

## BML

### Scrape tide data from http://tbone.biol.sc.edu/tide/
```{r}
bml_tide <- read_html("http://tide.arthroinfo.org/tideshow.cgi?tplotdir=horiz;gx=640;gy=240;caltype=ndp;type=mrare;interval=00%3A15;glen=150;units=feet;year=2021;month=05;day=31;hour=09;min=30;tzone=local;d_year=;d_month=01;d_day=01;d_hour=00;d_min=00;ampm24=24;site=Bodega%20Harbor%20entrance%2C%20California") %>%
  html_elements("pre") %>% #select only the date, time, and tide values from the webpage
  html_text2() %>% #convert list to data table
  data.frame() %>% #convert table to data frame
  mutate(date_tide = str_split(., pattern = "\n")) %>% #split into rows by each time point
  unnest(date_tide) %>% #unnest into two columns 
  mutate(date_tide=as.factor(date_tide)) %>% #make column values factors 
  separate(date_tide, into = c("date", "space", "time", "time_zone", "tide"), sep="\\s") %>% #separate the values (separated by spaces) into their own columns
  select(-"space", -".") %>% #remove the "space" (blank space) column and duplicated column created by unnest() 
  unite("time", "time", "time_zone", sep="\ ",) %>% #join together time and time zone
  unite("date_time", "date", "time", sep="\ ") %>% #join together date and time/time zone
  drop_na() %>% #remove final row with NA (not sure why that's even there)
  mutate(date_time=ymd_hm(date_time), #apply lubridate to date/time column
         tide=as.numeric(tide)) #coerce tide values from character to numeric
```

### Extract temp values from HOBO logger, clean up for overlap with BML readings
```{r}
bml_temp <- read_csv(here("data","hobo_bml.csv")) %>% 
  filter(row_number()!=1) %>% # remove the first row
  clean_names() %>%
  dplyr::rename(timestamp=x2,
         temp_hobo_f=x3) %>%
  mutate(date_time=ymd_hms(timestamp), #aapply lubridate to date/time column
         date=format(date_time, '%m/%d/%Y'), #create only date column
         time=format(date_time, '%H:%M:%S'), #create only time column
         temp_hobo_c=((as.numeric(temp_hobo_f)-32)*(5/9))) %>% #convert F to C from HOBO temp data
  select(temp_hobo_c,date_time) %>% #select the only columns that really matter
  filter(date_time <= ymd_hms("2021-10-08 19:00:00"),
         date_time > ymd_hms("2021-06-10 06:15:00")) #remove dates where HOBO was out of field (and first observation when HOBO was deployed, temp is WAY high)
```
  
### Combine the tides with the pH/temp values for BML, plot it up!
```{r}
bml <- all %>%
  filter(site=="Bodega Bay")

bml_all_tide <- full_join(bml, bml_tide) %>%
  drop_na(c(tide, p_h)) %>% #whoops, started off collecting data every 10 minutes and then switched to 15 minutes (so account for that by removing pH and tide values that don't overlap)
  distinct(date_time, .keep_all = TRUE) #remove duplicated data (not sure where those came from)

bml_all <- full_join(bml_all_tide, bml_temp) %>%
  drop_na(c(tide, temp_hobo_c)) %>% #remove hobo measurements taken outside of Durafet measurements
  dplyr::rename(temp_durafet_c=temp_c) %>% #dplyr::rename column of temp values taken from Durafet
  mutate(temp_hobo_c=ifelse((date_time==ymd_hms("2021-06-11 08:45:00")), NA, temp_hobo_c)) %>% #some HOBO values are quite odd, replace with NA
  mutate(temp_c=ifelse(is.na(temp_hobo_c==TRUE), temp_durafet_c, temp_hobo_c)) #the HOBO was deployed weeks after the Durafet, so create a new column accounting for this

bml3 <- bml_all %>%
  pivot_longer(cols=temp_durafet_c:temp_c,
               names_to = "data",
               values_to = "value")
  
ggplot(bml3, aes(x=date_time, y=value, group=data)) +
  geom_line(aes(color=data), size=0.7) +
  scale_x_datetime(breaks = scales::date_breaks("1 week"), 
                    labels = date_format("%m/%d %H:%m")) +
  xlab("Date time") +
  theme_bw() +
  theme(axis.text.x=element_text(angle=90))

#ggsave(here("figures", "bml_pH_temp_tide.png"), height=20, width=40, units="cm")
```

### Remove measurements where the sensor was out of the water (initially a tide lower than -0.5, but the "heatwave" on 6/15/21 shows how the temperature spikes disappear after filtering for tide of 0.2)
```{r}
bml_detide <- bml_all %>%
  mutate(tide_height=tide,
         tide=ifelse(tide_height>=0.2, "high", "low")) %>%
  drop_na(c(p_h)) #drop observations that don't overlap (10 min vs 15 min sampling interval)

ggplot(bml_detide, aes(x=date_time)) +
  geom_line(aes(y=tide_height), color="red") + 
  geom_line(aes(y=temp_c), color="blue") + # Divide by 10 to get the same range than the temperature
  scale_x_datetime(breaks = scales::date_breaks("1 week"), 
                    labels = date_format("%m/%d %H:%m")) +
  xlab("Date time") +
  theme_bw() +
  theme(axis.text.x=element_text(angle=90))
```

### Let's filter more, now based on anomalous temperature measurements (spikes in July)
```{r}
bml_detide_temp <- bml_detide %>%
  filter(!(date_time >= ymd_hms("2021-07-10 06:00:00") & date_time <= ymd_hms("2021-07-10 07:30:00")),
         !(date_time >= ymd_hms("2021-07-21 06:15:00") & date_time <= ymd_hms("2021-07-21 07:15:00")),
         !(date_time >= ymd_hms("2021-07-22 06:15:00") & date_time <= ymd_hms("2021-07-22 08:00:00")),
         !(date_time >= ymd_hms("2021-07-23 07:15:00") & date_time <= ymd_hms("2021-07-23 08:30:00")))

ggplot(bml_detide_temp, aes(x=date_time)) +
  geom_line(aes(y=temp_c), color="red") + 
  geom_line(aes(y=tide_height), color="blue") + # Divide by 10 to get the same range than the temperature
  scale_x_datetime(breaks = scales::date_breaks("1 day"), 
                    labels = date_format("%m/%d %H:%m")) +
  xlab("Date time") +
  theme_bw() +
  theme(axis.text.x=element_text(angle=90))
```

## LOL 

### Scrape tide data from http://tbone.biol.sc.edu/tide/
```{r}
lol_tide <- read_html("http:///tide.arthroinfo.org/tideshow.cgi?tplotdir=horiz;gx=640;gy=240;caltype=ndp;type=mrare;interval=00%3A15;glen=150;units=feet;year=2021;month=06;day=14;hour=08;min=00;tzone=local;d_year=;d_month=01;d_day=01;d_hour=00;d_min=00;ampm24=24;site=Point%20Arguello%2C%20California") %>%
  html_elements("pre") %>% #select only the date, time, and tide values from the webpage
  html_text2() %>% #convert list to data table
  data.frame() %>% #convert table to data frame
  mutate(date_tide = str_split(., pattern = "\n")) %>% #split into rows by each time point
  unnest(date_tide) %>% #unnest into two columns 
  mutate(date_tide=as.factor(date_tide)) %>% #make column values factors 
  separate(date_tide, into = c("date", "space", "time", "time_zone", "tide"), sep="\\s") %>% #separate the values (separated by spaces) into their own columns
  select(-"space", -".") %>% #remove the "space" (blank space) column and duplicated column created by unnest() 
  unite("time", "time", "time_zone", sep="\ ",) %>% #join together time and time zone
  unite("date_time", "date", "time", sep="\ ") %>% #join together date and time/time zone
  drop_na() %>% #remove final row with NA (not sure why that's even there)
  mutate(date_time=ymd_hm(date_time), #apply lubridate to date/time column
         tide=as.numeric(tide)) #coerce tide values from character to numeric
```

### Extract temp values from HOBO logger, clean up for overlap with LOL readings
```{r}
lol_temp <- read_csv(here("data","hobo_lol.csv")) %>% 
  filter(row_number()!=1) %>% # remove the first row
  clean_names() %>%
  dplyr::rename(timestamp=x2,
         temp_hobo_f=x3) %>%
  mutate(date_time=ymd_hms(timestamp), #aapply lubridate to date/time column
         date=format(date_time, '%m/%d/%Y'), #create only date column
         time=format(date_time, '%H:%M:%S'), #create only time column
         temp_hobo_c=((as.numeric(temp_hobo_f)-32)*(5/9))) %>% #convert F to C from HOBO temp data
  select(temp_hobo_c,date_time) #select the only columns that really matter
```

### Combine the tides with the pH/temp values for LOL, plot it up!
```{r}
# filter all data for LOL
lol <- all %>%
  filter(site=="Lompoc Landing") %>%
  distinct(date_time, .keep_all = TRUE) #remove duplicated data (not sure where those came from)

#join HOBO data with Durafet data
lol_all_1 <- full_join(lol, lol_temp) %>%
  drop_na(c(temp_c, p_h)) %>% #remove hobo measurements taken outside of Durafet measurements
  dplyr::rename(temp_durafet_c=temp_c) 

#Check difference between Durafet and HOBO temperature measurements
lol_temp_diff <- lol_all_1 %>%
  mutate(diff=temp_durafet_c-temp_hobo_c)

#Plot it up to check for anomalies, ID where sensor was out of water
y1 <- lol_temp_diff$temp_durafet_c
y2 <- lol_temp_diff$temp_hobo_c
y3 <- lol_temp_diff$diff
x <- lol_temp_diff$date_time

highchart() %>% 
  hc_add_series(data = y1, dashStyle="solid") %>% 
  hc_add_series(data = y2, yAxis = 1) %>% 
  hc_add_series(data = y3, yAxis = 2) %>%
  hc_yAxis_multiples(
     list(lineWidth = 3, lineColor='#D55E00', title=list(text="durafet")),
     list(lineWidth = 3, lineColor="#009E73", title=list(text="hobo")),
     list(lineWidth = 3, lineColor="#0072B2", title=list(text="diff"))) %>%
    hc_xAxis(title = "Date", categories = x, breaks=10) %>%
  hc_colors(c("#D55E00","#009E73", "#0072B2"))

#Join pH, temp data with tide data
lol_all <- full_join(lol_all_1, lol_tide) %>%
  drop_na(c(tide, p_h)) %>% #remove unnecessary tide values
  dplyr::rename(temp_c=temp_hobo_c) 
```

### Use wave data to check tide vs temperature
```{r}
#look at July NOAA NDBC data from Santa Maria buoy 
lol_wave <- read_csv(here("data","wave-july2021.csv")) %>% 
  clean_names() %>%
  filter(wvht<99.00) %>%
  mutate(mn=mn+5) %>%
  unite("date", "year", "month", "day", sep="/") %>%
  unite("time", "hr", "mn", sep=":") %>%
  unite("date_time", "date", "time", sep=" ") %>%
  mutate(date_time=ymd_hm(date_time)) #apply lubridate to date and time

#join this with the july sensor df
lol_july <- lol_all %>%
  filter(date_time < ymd_hms("2021-08-01 00:00:00"),
         date_time > ymd_hms("2021-07-01 00:00:00"))

lol_wave_join <- full_join(lol_july, lol_wave) %>%
  fill(wvht, .direction = "updown") %>% #fill in empty wave measurements with previous values before
  distinct(date_time, .keep_all = TRUE) %>% #remove duplicated data (not sure where those came from)
  filter(date_time > ymd_hms("2021-07-20 00:00:00"),
         date_time < ymd_hms("2021-07-25 00:00:00"))

y1 <- lol_wave_join$temp_c
y2 <- lol_wave_join$wvht
y3 <- lol_wave_join$tide
x <- lol_wave_join$date_time

highchart() %>% 
  hc_add_series(data = y1, dashStyle="solid") %>% 
  hc_add_series(data = y2, yAxis = 1) %>% 
  hc_add_series(data = y3, yAxis = 1) %>%
  hc_yAxis_multiples(
     list(lineWidth = 3, lineColor='#D55E00', title=list(text="Temp")),
     list(lineWidth = 3, lineColor="#009E73", title=list(text="Wave Height")),
     list(lineWidth = 3, lineColor="#0072B2", title=list(text="Tide cycle"))) %>%
    hc_xAxis(title = "Date", categories = x, breaks=10) %>%
  hc_colors(c("#D55E00",
              "#009E73",
              "#0072B2"))
```
The wave data isn't a magic bullet, but it does look like I might be overprocessing the data if I make the tide height as high as 3.0. Based on the July data (especially keeping in mind temperature changes will be most apparent in the middle of the day when the sun is out in full force), I'm going to remove any data below a tide height of 2.0

### Let's filter based on anomalous pH measurements (and remove measurements where the sensor pool was disconnected from the ocean)
```{r}
lol_detided <- lol_all %>%
  mutate(tide_height=tide,
         tide=ifelse(tide_height>=3.232, "high", "low")) %>%
  filter(tide=="high") %>%
  filter(!(date_time >= ymd_hms("2021-07-31 13:30:00") & date_time <= ymd_hms("2021-07-31 14:00:00")),
         (date_time != ymd_hms("2021-09-15 09:15:00")),
         !(date_time >= ymd_hms("2021-09-16 01:30:00") & date_time <= ymd_hms("2021-09-16 02:45:00")), #anomalous points here identified by comparing durafet and hobo temp logger values
         !(date_time >= ymd_hms("2021-08-01 13:45:00") & date_time < ymd_hms("2021-08-01 15:00:00"))) %>%
  drop_na() #remove times with tides but without pH values
```

## ALG 

### Scrape tide data from http://tbone.biol.sc.edu/tide/
```{r}
alg_tide <- read_html("http://tide.arthroinfo.org/tideshow.cgi?tplotdir=horiz;gx=640;gy=240;caltype=ndp;type=mrare;interval=00%3A15;glen=150;units=feet;year=2021;month=06;day=12;hour=01;min=00;tzone=local;d_year=;d_month=01;d_day=01;d_hour=00;d_min=00;ampm24=24;site=Gaviota%2C%20California") %>%
  html_elements("pre") %>% #select only the date, time, and tide values from the webpage
  html_text2() %>% #convert list to data table
  data.frame() %>% #convert table to data frame
  mutate(date_tide = str_split(., pattern = "\n")) %>% #split into rows by each time point
  unnest(date_tide) %>% #unnest into two columns 
  mutate(date_tide=as.factor(date_tide)) %>% #make column values factors 
  separate(date_tide, into = c("date", "space", "time", "time_zone", "tide"), sep="\\s") %>% #separate the values (separated by spaces) into their own columns
  select(-"space", -".") %>% #remove the "space" (blank space) column and duplicated column created by unnest() 
  unite("time", "time", "time_zone", sep="\ ",) %>% #join together time and time zone
  unite("date_time", "date", "time", sep="\ ") %>% #join together date and time/time zone
  drop_na() %>% #remove final row with NA (not sure why that's even there)
  mutate(date_time=ymd_hm(date_time), #apply lubridate to date/time column
         tide=as.numeric(tide)) #coerce tide values from character to numeric
```

### Combine the tides with the pH/temp values for ALG, plot it up!
```{r}
# filter all data for ALG
alg <- all %>%
  filter(site=="Alegria")

#combine tide data with sensor data
alg_all_1 <- full_join(alg, alg_tide) %>%
  drop_na(c(temp_c)) #remove extra tide measurements

#extract HOBO data
alg_temp <- read_csv(here("data","hobo_alg.csv")) %>% 
  filter(row_number()!=1) %>% # remove the first row
  clean_names() %>%
  dplyr::rename(timestamp=x2,
         temp_hobo_f=x3) %>%
  mutate(date_time=ymd_hms(timestamp), #aapply lubridate to date/time column
         date=format(date_time, '%m/%d/%Y'), #create only date column
         time=format(date_time, '%H:%M:%S'), #create only time column
         temp_hobo_c=((as.numeric(temp_hobo_f)-32)*(5/9))) %>% #convert F to C from HOBO temp data
  select(temp_hobo_c,date_time) #select the only columns that really matter

# Join HOBO data to alegria df
alg_all <- full_join(alg_all_1, alg_temp) %>%
  drop_na(c(temp_c)) %>% #remove hobo measurements taken outside of Durafet measurements
  dplyr::rename(temp_durafet_c=temp_c) 

alg_plot <- alg_all %>%
  pivot_longer(cols=temp_durafet_c:temp_hobo_c,
               names_to = "data",
               values_to = "value")
  
ggplot(alg_plot, aes(x=date_time, y=value, group=data)) +
  geom_line(aes(color=data), size=0.7) +
  scale_x_datetime(breaks = scales::date_breaks("1 week"), 
                    labels = date_format("%m/%d %H:%m")) +
  xlab("Date time") +
  theme_bw() +
  theme(axis.text.x=element_text(angle=90))

#ggsave(here("figures", "alg_pH_temp_tide.png"), height=20, width=40, units="cm")
```

### Remove anomalous measurements at beginning of deployment (sensor was likely inundated with sand)
```{r}
alg_desand <- alg_all %>%
  filter(date_time > ymd_hms("2021-06-18 01:00:00")) %>% #super weird, but pH looks like it stabilizes at 1AM post sand
  distinct(date_time, .keep_all = TRUE) #remove duplicated data (not sure where those came from)

ggplot(alg_desand, aes(x=date_time, y=p_h)) +
  geom_line(size=0.7) +
  scale_x_datetime(breaks = scales::date_breaks("1 week"), 
                    labels = date_format("%m/%d %H:%m")) +
  xlab("Date time") +
  theme_bw() +
  theme(axis.text.x=element_text(angle=90))
```
There's still some noise, let's filter those dates out of the data

### Remove (or smooth?) time periods with a LOT of noise, and de-tide data
```{r}
alg_detide <- alg_desand %>%
  mutate(tide_height=tide,
         tide=ifelse(tide_height>=0.9, "high", "low")) %>%
  filter(tide=="high") %>%
  mutate(p_h=ifelse((date_time > ymd_hms("2021-10-25 19:00:00") & date_time < ymd_hms("2021-11-02 10:45:00")), NA, p_h),
         temp_hobo_c=ifelse((date_time > ymd_hms("2021-10-25 19:00:00") & date_time < ymd_hms("2021-11-02 10:45:00")), NA, temp_hobo_c)) %>% #since this is in the middle of the dataset, replace values with "NA" rather than remove entirely (otherwise when plotting, R will fill in the missing space with a straight line)
  #filter(!(date_time > ymd_hms("2021-10-25 19:00:00") & date_time < ymd_hms("2021-11-02 10:45:00"))) %>% #looks like another time when the sensor was sanded
  mutate(p_h=ifelse((date_time > ymd_hms("2021-07-17 11:00:00") & date_time < ymd_hms("2021-07-17 19:30:00")), NA, p_h),
         temp_hobo_c=ifelse((date_time > ymd_hms("2021-07-17 11:00:00") & date_time < ymd_hms("2021-07-17 19:30:00")), NA, temp_hobo_c)) %>%
    # filter(!(date_time > ymd_hms("2021-07-17 11:00:00") & date_time < ymd_hms("2021-07-17 19:30:00"))) %>% #this is a chunk of time with a lot of noise
  mutate(p_h=ifelse((date_time > ymd_hms("2021-09-09 10:15:00") & date_time < ymd_hms("2021-09-09 13:00:00")), NA, p_h),
         temp_hobo_c=ifelse((date_time > ymd_hms("2021-09-09 10:15:00") & date_time < ymd_hms("2021-09-09 13:00:00")), NA, temp_hobo_c)) %>%
  #filter(!(date_time > ymd_hms("2021-09-09 10:15:00") & date_time < ymd_hms("2021-09-09 13:00:00"))) %>% #this is a chunk of time with a lot of noise
  filter(!(date_time %in% (ymd_hms("2021-06-19 10:15:00", #remove individual time points with significant jumps in the wrong direction, identified using highchart
                                  "2021-06-29 14:15:00",
                                  "2021-06-29 15:30:00",
                                  "2021-06-30 11:45:00",
                                  "2021-06-30 14:15:00",
                                  "2021-06-30 16:15:00",
                                  "2021-06-30 19:15:00",
                                  "2021-07-02 06:15:00",
                                  "2021-07-02 07:45:00",
                                  "2021-07-02 10:45:00",
                                  "2021-07-04 09:45:00",
                                  "2021-07-04 12:00:00",
                                  "2021-07-04 15:15:00",
                                  "2021-10-02 18:15:00",
                                  "2021-10-09 11:15:00",
                                  "2021-10-22 02:30:00",
                                  "2021-10-22 08:15:00",
                                  "2021-10-23 16:30:00",
                                  "2021-10-25 06:45:00",
                                  "2021-10-25 14:45:00",
                                  "2021-10-25 15:00:00",
                                  "2021-10-25 16:30:00"

)))) %>%
  dplyr::rename(temp_c=temp_hobo_c) #dplyr::rename this column so I can merge ALG with all other sites

ggplot(alg_detide, aes(x=date_time)) +
  geom_line(aes(y=p_h), color="#009E73") + 
  geom_line(aes(y=temp_c), color="#D55E00") + # Divide by 10 to get the same range than the temperature
  geom_line(aes(y=tide_height), color="#0072B2") + # Divide by 10 to get the same range than the temperature
  scale_x_datetime(breaks = scales::date_breaks("4 days"), 
                    labels = date_format("%m/%d %H:%m")) +
  annotate(geom="text", x=as.POSIXct("2021-6-18 00:01:00"), y=19, hjust=0, label="Temp (C)", color="#D55E00", size=5) +
  annotate(geom="text", x=as.POSIXct("2021-6-18 00:01:00"), y=10, hjust=-0.1, label="pH", color="#009E73", size=5) +
  annotate(geom="text", x=as.POSIXct("2021-6-18 00:01:00"), y=-1.5, hjust=0, label="Tide", color="#0072B2", size=5) +
  xlab("Date & time") +
  ylab("Value") +
  theme_bw() +
  theme(axis.text.x=element_text(angle=45, vjust = 1, hjust=1, size=12),
        axis.title.x=element_text(size=15),
        axis.text.y=element_text(size=12),
        axis.title.y=element_text(size=15))
```

# Now let's actually plot up a comparison of the pH values across sites
```{r}
#join all sites into one df (and remove diff columns)
ph_all <- merge_recurse(list(alg_detide, lol_detided, bml_detide_temp))

#set order for sites
ph_all$site <- factor(ph_all$site, levels=c("Lompoc Landing","Alegria","Bodega Bay"))

#set custom color for sites
pal <- c(
  "Alegria" = "#D55E00",
  "Lompoc Landing" = "#009E73", 
  "Bodega Bay" = "#0072B2"
)

ggplot(ph_all, aes(x=date_time, y=p_h, group=site)) +
  geom_line(aes(color=site), size=0.7, alpha=0.8) +
  scale_color_manual(values = pal) + #color lines by custom site color palette
  scale_x_datetime(breaks = scales::date_breaks("1 week"), 
                    labels = date_format("%m/%d %H:%m")) +
  #xlab("Date and time") +
  ylab("pH") +
  theme_bw() +
  theme(axis.text.x=element_blank(), 
        #element_text(angle=45, vjust = 1, hjust=1, size=12),
        axis.title.x=element_blank(),
        #element_text(size=15),
        axis.text.y=element_text(size=12),
        #axis.ticks.x=element_blank(),
        axis.title.y=element_text(size=15),
        legend.position = "none")

ggsave(here("figures", "ph_all.png"), height=20, width=40, units="cm")

# For presentations, highlight one line at a time per site
#BML focal
ph_all$site <- factor(ph_all$site, levels=c("Lompoc Landing","Alegria","Bodega Bay"))

ggplot(ph_all, aes(x=date_time, y=p_h, group=site)) +
  geom_line(aes(color=site, alpha=site), size=0.7) +
  scale_color_manual(values = pal) + #color lines by custom site color palette
  scale_x_datetime(breaks = scales::date_breaks("1 week"), 
                    labels = date_format("%m/%d %H:%m")) +
  xlab("Date time") +
  ylab("pH") +
  theme_bw() +
  theme(axis.text.x=element_text(angle=45, vjust = 1, hjust=1, size=12),
        axis.title.x=element_text(size=15),
        axis.text.y=element_text(size=12),
        axis.title.y=element_text(size=15),
        legend.position = "none") +
  scale_alpha_manual(values=c(0.5,0.5,1))

ggsave(here("figures", "ph_bml.png"), height=20, width=40, units="cm")

#LOL focal
ph_all$site <- factor(ph_all$site, levels=c("Alegria","Bodega Bay", "Lompoc Landing"))

ggplot(ph_all, aes(x=date_time, y=p_h, group=site)) +
  geom_line(aes(color=site, alpha=site), size=0.7) +
  scale_color_manual(values = pal) + #color lines by custom site color palette
  scale_x_datetime(breaks = scales::date_breaks("1 week"), 
                    labels = date_format("%m/%d %H:%m")) +
  xlab("Date time") +
  ylab("pH") +
  theme_bw() +
  theme(axis.text.x=element_text(angle=45, vjust = 1, hjust=1, size=12),
        axis.title.x=element_text(size=15),
        axis.text.y=element_text(size=12),
        axis.title.y=element_text(size=15),
        legend.position = "none") +
  scale_alpha_manual(values=c(0.5,0.5,1))

ggsave(here("figures", "ph_lol.png"), height=20, width=40, units="cm")

#ALG focal
ph_all$site <- factor(ph_all$site, levels=c("Lompoc Landing","Bodega Bay", "Alegria"))

ggplot(ph_all, aes(x=date_time, y=p_h, group=site)) +
  geom_line(aes(color=site, alpha=site), size=0.7) +
  scale_color_manual(values = pal) + #color lines by custom site color palette
  scale_x_datetime(breaks = scales::date_breaks("1 week"), 
                    labels = date_format("%m/%d %H:%m")) +
  xlab("Date time") +
  ylab("pH") +
  theme_bw() +
  theme(axis.text.x=element_text(angle=45, vjust = 1, hjust=1, size=12),
        axis.title.x=element_text(size=15),
        axis.text.y=element_text(size=12),
        axis.title.y=element_text(size=15),
        legend.position = "none") +
  scale_alpha_manual(values=c(0.5,0.5,1))

ggsave(here("figures", "ph_alg.png"), height=20, width=40, units="cm")

# Save the "final" df to a .csv file - but do this before you "detide" all the data
#write.csv(ph_all, here("data", "ph_clean.csv"), row.names = FALSE)
```

# Let's also compare temperature across sites
```{r}
#add site identity before joining
bml_temp_join <- bml_detide_temp %>%
  select(date_time, temp_c, p_h) %>%
  mutate(site = "Bodega Bay")

lol_temp_join <- lol_detided %>%
  select(date_time, temp_c, p_h) %>%
  mutate(site = "Lompoc Landing")

alg_join <- alg_detide %>%
  select(date_time, temp_c, p_h) %>%
  mutate(site = "Alegria")

#join these all together into one df
temp_all <- merge_recurse(list(bml_temp_join, lol_temp_join, alg_join))

#set order for sites
temp_all$site <- factor(temp_all$site, levels=c("Lompoc Landing","Alegria","Bodega Bay"))

#set custom color for sites
pal <- c(
  "Alegria" = "#D55E00",
  "Lompoc Landing" = "#009E73", 
  "Bodega Bay" = "#0072B2"
)

ggplot(temp_all, aes(x=date_time, y=temp_c, group=site)) +
  geom_line(aes(color=site), size=0.7, alpha=0.8) +
  scale_color_manual(values = pal) + #color lines by custom site color palette
  scale_x_datetime(breaks = scales::date_breaks("1 week"), 
                    labels = date_format("%m/%d %H:%m")) +
  xlab("Date and time") +
  ylab("Temperature") +
  theme_bw() +
  theme(axis.text.x=element_text(angle=45, vjust = 1, hjust=1, size=12),
        axis.title.x=element_text(size=15),
        axis.text.y=element_text(size=12),
        axis.title.y=element_text(size=15),
        legend.position = "none")

ggsave(here("figures", "temp_all.png"), height=20, width=40, units="cm")
```

## Let's try to combine pH and temperature plots into one grid

```{r}
all <- temp_all %>%
  relocate(site, date_time, temp_c, p_h) %>%
  dplyr::rename(Site=site) %>%
  pivot_longer(cols=temp_c:p_h,
               names_to = "group",
               values_to = "value")

#set order for sites
all$Site <- factor(all$Site, levels=c("Lompoc Landing","Alegria","Bodega Bay"))

#set custom color for sites
pal <- c(
  "Alegria" = "#D55E00",
  "Lompoc Landing" = "#009E73", 
  "Bodega Bay" = "#0072B2"
)

ggplot(all, aes(x=date_time, y=value, group=group)) +
  geom_line(aes(color=Site), size=0.7, alpha=0.7) +
  scale_color_manual(values = pal) + #color lines by custom site color palette
  #geom_point(aes(color=group), size=0.5) +
  scale_x_datetime(breaks = scales::date_breaks("1 week"), 
                    labels = date_format("%m/%d %H:%m")) +
  facet_grid(group ~ ., #facet wrap to create one panel for pH and one for temp
             scales = "free_y",
             switch="both",
             labeller = as_labeller(c(temp_c = "Temperature (C)", p_h = "pH"))) + #customize strip labels
  xlab("Date and time") +
  ylab(NULL) + #remove "Value" from Y axis label
  theme_bw() +
  theme(axis.text.x=element_text(angle=45, vjust = 1, hjust=1, size=22),
        axis.title.x=element_text(size=30),
        axis.text.y=element_text(size=22),
        legend.text=element_text(size=22),
        legend.title = element_blank(),
        #element_text(size=25),
        legend.position = c(0.92,0.95), #customize legend position on plot
        #legend.key.size = unit(4,"line"),
        #legend.key.height = unit(1,"cm"),
        strip.background = element_blank(), #remove strip background from facet_grid
        strip.text.y = element_text(size = 30),
        strip.placement = "outside") + #place the strip outside of the plot
  guides(colour = guide_legend(override.aes = list(size=5, alpha = 1)))

ggsave(here("figures", "temp_all_facet.png"), height=50, width=60, units="cm")
```

Now, for some "Chan et al 2017" style analyses

## Look at the frequency of pH values below 7.8 
```{r}
ph_7.8 <- ph_all %>%
  #drop_na(c(tide)) %>%
  mutate(threshold = ifelse(p_h<7.8, "below", "above")) %>%
  filter(!is.na(p_h)) %>% #remove the times with anomalous pH readings that were replaced with NA
  group_by(site, threshold) %>%
  summarize(tot_observations = n()) %>%
  ungroup() %>%
  add_row(site="Alegria", threshold="below", tot_observations=0) #add this row because Alegria has 0 observations below 7.8 (for now...)

#Get total number of observations - use double square brackets because it's a tibble
lol_n <- ph_7.8[[1,3]] + ph_7.8[[2,3]] 
bml_n <- ph_7.8[[3,3]] + ph_7.8[[4,3]]
alg_n <- ph_7.8[[5,3]] + ph_7.8[[6,3]]

#Get frequency of pH values below 7.8
lol_7.8 <- ph_7.8[[2,3]]/lol_n
bml_7.8 <- ph_7.8[[4,3]]/bml_n
alg_7.8 <- ph_7.8[[6,3]]/alg_n
```

# Create a nice table
```{r}
#create df for ALG without NA values
alg_no_na <- alg_detide %>%
  filter(!is.na(p_h))
  
# Calculate min, max, average pH values for each site
alg_min <- (min(alg_no_na$p_h))
alg_max <- (max(alg_no_na$p_h))
alg_median <- (median(alg_no_na$p_h))
alg_mean <- (mean(alg_no_na$p_h))
alg_sd <- (sd(alg_no_na$p_h))
alg_cv <- (alg_sd/alg_mean)*100 #coefficient of variation

lol_min <- (min(lol_detided$p_h))
lol_max <- (max(lol_detided$p_h))
lol_median <- (median(lol_detided$p_h))
lol_mean <- (mean(lol_detided$p_h))
lol_sd <- (sd(lol_detided$p_h))
lol_cv <- (lol_sd/lol_mean)*100 #coefficient of variation

bml_min <- (min(bml_detide_temp$p_h))
bml_max <- (max(bml_detide_temp$p_h))
bml_median <- (median(bml_detide_temp$p_h))
bml_mean <- (mean(bml_detide_temp$p_h))
bml_sd <- (sd(bml_detide_temp$p_h))
bml_cv <- (bml_sd/bml_mean)*100 #coefficient of variation

ph_table <- tribble(
  ~"Site", ~"Min pH", ~"Max pH", ~"Median pH",  ~"CV", ~"< 7.8", 
  "Bodega Marine Lab", bml_min, bml_max, bml_median, bml_cv, bml_7.8,
  "Lompoc Landing", lol_min, lol_max, lol_median, lol_cv, lol_7.8,
  "Alegria", alg_min, alg_max, alg_median, alg_cv, alg_7.8) 

ph_table %>%
  gt() %>%
  fmt_number(
  columns = c("Min pH":"Median pH"),
  rows = everything(),
  decimals = 2) %>%
  data_color(
    columns = c("Site"),
    colors = scales::col_factor( # <- bc it's a factor
      palette = c("#D55E00","#0072B2","#009E73"),
      domain = c("Bodega Marine Lab", "Lompoc Landing", "Alegria"))) %>%
  gtsave(here("figures", "table_ph.png"))

# Calculate min, max, average temp values for each site
alg_min <- (min(alg_no_na$temp_c))
alg_max <- (max(alg_no_na$temp_c))
alg_median <- (median(alg_no_na$temp_c))
alg_mean <- (mean(alg_no_na$temp_c))
alg_sd <- (sd(alg_no_na$temp_c))
alg_cv <- (alg_sd/alg_mean)*100 #coefficient of variation

lol_min <- (min(lol_detided$temp_c))
lol_max <- (max(lol_detided$temp_c))
lol_median <- (median(lol_detided$temp_c))
lol_mean <- (mean(lol_detided$temp_c))
lol_sd <- (sd(lol_detided$temp_c))
lol_cv <- (lol_sd/lol_mean)*100 #coefficient of variation

bml_min <- (min(bml_detide_temp$temp_c))
bml_max <- (max(bml_detide_temp$temp_c))
bml_median <- (median(bml_detide_temp$temp_c))
bml_mean <- (mean(bml_detide_temp$temp_c))
bml_sd <- (sd(bml_detide_temp$temp_c))
bml_cv <- (bml_sd/bml_mean)*100 #coefficient of variation

temp_table <- tribble(
  ~"Site", ~"Min temp", ~"Max temp", ~"Median temp",  ~"CV", 
  "Bodega Marine Lab", bml_min, bml_max, bml_median, bml_cv,
  "Lompoc Landing", lol_min, lol_max, lol_median, lol_cv,
  "Alegria", alg_min, alg_max, alg_median, alg_cv) 

temp_table %>%
  gt() %>%
  fmt_number(
  columns = c("Min temp":"Median temp"),
  rows = everything(),
  decimals = 2) %>%
  data_color(
    columns = c("Site"),
    colors = scales::col_factor( # <- bc it's a factor
      palette = c("#D55E00","#0072B2","#009E73"),
      domain = c("Bodega Marine Lab", "Lompoc Landing", "Alegria"))) %>%
  gtsave(here("figures", "table_temp.png"))
```

# Summary (what I did)

At BML, sampling interval started at every 10 minutes and then switched to 15 minutes for all sites

Filtering for tide heights based on temp/pH patterns
- All points at ALG were removed at tide heights below 0.9
- All points at LOL were removed at tide heights below 2.0 (3.232 was also a good option, but I don't want to remove more values than need be... ask about this) 
- All points at BML were removed at tide heights below 0.2

Removed additional anomalous time points (noise or major jumps, especially in wrong direction)
- 20 from BML
- All data collected before 06/18/2021 at ALG, data collected between 10/25/2021 and 11/02/2021, plus 65 extra points
- 9 from LOL

# Next: zoom in on LOL as an example of pH variation within tidepool
### Create df (remove anomalous points) - use this for "lol-analysis.Rmd"
```{r}
lol_cycle <- lol_detided %>%
  select(-date, -time, -site) %>%
  relocate(date_time, sensor_number, tide_height, tide, temp_c, temp_durafet_c, p_h)

# Save the "final" df to a .csv file
#write.csv(lol_cycle, here("data", "lol_all.csv"), row.names = FALSE)
```


Next steps:
1. Talk to someone about overfiltering vs underfiltering LOL data - 3.232 tide vs 2.0 tide
2. Triple check filtering of all sites (for pH and temp anomalies) so data can be published